Molecular dynamics simulation of synchronization in driven particles

### 2020-12-18 manuscript/manuscript_ajp_2020.tex
rearranging the intro to be more undergrad compelling
liking the reshape!

\documentclass[prb,preprint]{revtex4} 
% The line above defines the type of LaTeX document.
% Note that AJP uses the same style as Phys. Rev. B (prb).

@@

\begin{abstract}

  %DM editorial - too many ``we'' statements, not enough emphasize on importance.
  %more detail on the methods?
  %will need to rewrite results/conclusions as we shape the manuscript.

  %What problem did you study and why is it important?
  We study synchronization of interacting particles, 
  where the particles are 
  confined to a narrow channel
  and 
  driven by an external applied force.  
  %What methods did you use?
  Our
  numerical simulations
  mimick experimental studies of driven colloidal particles
  confined by light-fields.
  Much like experiments on colloids, 
  we control the particle interactions,
  external force and confining environment.
  We assume the particles move through
  a viscous liquid,
  so we model the dynamics of each particle
  with 
  overdamped equations of motion.
  %What were your principal results?
  We observe
  particle synchronization under a variety of conditions,
  including static and dynamic patterns formed on
  periodic and quasiperiod substrates
  and the propagation of 
  high density regions of particles,
  or kinks.
  %What conclusions have you drawn from these results?
  We demonstrate the transition from
  trapped to sliding dynamics in
  quasiperiodic landscapes differs
  from that of a periodic landscape by [TBD],
  and that kinks of high density propagate
  [more/less ?  ] when [what?].
  We also explore 
  the limits of applied force that
  cause the dynamics to transition to chaotic behavior.
  We include sample code
  and exercises for students
  %that include
  with 
  opportunties
  to reproduce our results and propose
  new numerical experiments.
  With only a few particles in two-dimensions,
  the simulation runs quickly,
  making this an appropriate model for undergraduates to explore.

@@

\section{Introduction} %
%
%Outline
%------------------------------------------------------------------
%synchronization = widest audience possible
Synchronization is a universal phenomena
that can be modeled with numerical studies of confined 
driven particles.
%Christiaan Huygens 
%[has been adapted into many useful technologies...]
Examples of synchronization occur
in many biological systems with periodic behavior,
from fireflies signaling in time with a pulsed light emitting diode \cite{},
to coordinated flapping of birds wings \cite{},
to simultaneous firing of neurons in cardiac muscle \cite{}
and brain tissue \cite{}.

Huygens 1665 observation of
in-phase oscillation of pendulum clocks \cite{Pikovsky2003},
to

@@

%single DC case with figure
%dc_current 0.01
%ac_current 0.2
%ac_frequency 0.0001
%fp_y 0.1
\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[scale=.25]{position}
\caption{The position as a function of time of a single driven particle
  normalized by the period of the substrate $\lambda$.
  }
\label{fig:0}
\end{figure}
\end{center}

@@

  \bibitem{} Reichhardt, C., & Reichhardt, C. J. O. (2015). Shapiro steps for skyrmion motion on a washboard potential with longitudinal and transverse ac drives. Physical Review B - Condensed Matter and Materials Physics, 92(22). https://doi.org/10.1103/PhysRevB.92.224432

>>>>

%\documentclass[prb,preprint]{revtex4} 
\documentclass[twocolumn,preprintnumbers,amsmath,amssymb,aps,prx]{revtex4}
% The line above defines the type of LaTeX document.
% Note that AJP uses the same style as Phys. Rev. B (prb).

@@

\begin{abstract}
  %What problem did you study and why is it important?
  We discuss a
  numerical model
  of particles
  confined to a narrow channel 
  driven
  across a washboard potential energy landscape.
  The model 
  exhibits
  synchronization effects that provide insight
  to the behavior of 
  many experimental systems of
  interest to condensed matter physics 
  including 
  magnetically driven colloidal particles
  confined by light-fields, 
  voltage driven superconducting vortices
  confined in a Josephson junction,
  and ac and dc driven
  charge and spin density waves.
  We present the basics
  of the molecular dynamics simulations
  for particles moving through
  a viscous liquid.
  %so we model the dynamics of each particle
  %with 
  %overdamped equations of motion.
  We demonstrate how to visualize
  the system
  and measure hallmarks of synchronization, 
  Shapiro steps,
  that exhibit non-Ohmic step-like behavior 
  in the 
  the particle velocity as a function of applied driving force.
  We include sample code
  and exercises for students
  %that include
  with 
  opportunties
  to reproduce our results and propose
  new numerical experiments.
  With only a few particles in two-dimensions,
  the simulation runs quickly,
  making this an appropriate model for undergraduates to explore.

@@

\section{Introduction} %
%
%Outline
%------------------------------------------------------------------
%synchronization = widest audience possible
Synchronization is a universal phenomena
where individual oscillators adjust rhythm due
to an external stimuli \cite{Pikovsky2003}.
%Christiaan Huygens 
%[has been adapted into many useful technologies...]
Huygens' 1665 experiments on
the 
motions of two pendula on wall-mounted clocks,
demonstrated the tendency of 
two periodic oscillators in close
proximity to swing in time, i.e. to synchronize \cite{Bennett2002}.
Physical synchronous behavior are observed
in many everyday systems
such as the
coupled flickering candle flames \cite{Okamoto2016} 
and metronomes \cite{Jia2015}.
Biological systems benefit from cooperative
synchronzation.
Birds coordinate
flapping of wings in order to optimize energy use during flight \cite{Portugal2014},
frogs communicate by croaking patterns dictated by location \cite{Aihara2014},
and humans synchronize clapping in time with music \cite{Tranchant2016}.

to simultaneous firing of neurons in cardiac muscle \cite{}
and brain tissue (i.e. brain waves) \cite{}.

Such systems can be driven to form new synchronized patterns,
for instance
firefly [swarms] signal in time with a pulsed light emitting diode \cite{},
[more examples]
Driven coupled oscillations are sufficiently
broad and useful to be the subject of computational modeling.

@@

%single DC case with figure
%dc_current 0.01
%ac_current 0.2
%ac_frequency 0.0001
%fp_y 0.1
\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{position}
\caption{The position as a function of time of a single driven particle
  normalized by the period of the substrate $\lambda$.
  }
\label{fig:0}
\end{figure}
\end{center}

@@

\bibitem{Bennett2002} M. Bennett, M.F. Schatz, H. Rockwood, and K. Wiesenfeld, Huygen's clocks, Proc. Roy. Soc. A {\bf 458}, 563 (2002).

\bibitem{Okamoto2016} K. Okamoto, A. Kijima, Y. Umeno, and H. Shima. Synchronization in flickering of three-coupled candle flames. Sci Rep {\bf 6}, 36145 (2016)

\bibitem{Jia2015}  J. Jia, Z. Song, W. Liu, J. Kurths, and Xiao, J. Experimental study of the triplet synchronization of coupled nonidentical mechanical metronomes. Sci. Rep. {\bf 5}, 17008 (2015).

\bibitem{Portugal2014} S. Portugal, T. Hubel, J. Fritz, S. Heese, D. Trobe, B. Voelkl, S. Hailes, A. M. Wilson and J. R. Usherwood.  Upwash exploitation and downwash avoidance by flap phasing in ibis formation flight. Nature {\bf 505}, 399 (2014).

  \bibitem{Aihara2014} I. Aihara, T. Mizumoto, T. Otsuka, H. Awano, K. Nagira, H. G. Okuno and K. Aihara. Spatio-Temporal Dynamics in Collective Frog Choruses Examined by Mathematical Modeling and Field Observations. Sci Rep {\bf 4}, 3891 (2014).

    %  \bibitem{} Reichhardt, C., & Reichhardt, C. J. O. (2015). Shapiro steps for skyrmion motion on a washboard potential with longitudinal and transverse ac drives. Physical Review B - Condensed Matter and Materials Physics, 92(22). https://doi.org/10.1103/PhysRevB.92.224432

    \bibitem{Tranchant2016} P. Tranchant, D. T. Vuvan, and I. Peretz, Keeping the Beat: A Large Sample Study of Bouncing and Clapping to Music. PLoS ONE 11(7): e0160178. (2016).
